Auditory Scene Similarity Retrieval and Classification with Relevance based Quantization of Scattering Features

### 2016-12-12 paper/structureScene.tex
Edit feature design section

Before classification or clustering, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
Whether we adopt a time-frequency perspective or a machine learning perspective on the problem of acoustic scene classification, such operations would either qualify as pre- or post-processing.

@@

Many algorithms in pattern recognition, including nearest-neighbor classifiers and support vector machines (SVM)s, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation, such as a logarithm, to all coefficients.
Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.

@@

Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone -- a source of spurious variability in scene classification.
Logarithmic compression can linearize this dependency, facilitating the construction of a powerful invariant at the classifier stage.

For the task of musical genre recognition, second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ are sometimes normalized by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, since this decorrelates them from one another \cite{Anden2014}.
We note that taking the logarithm of such renormalized coefficients yields
\begin{equation}
\log \dfrac{\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)}{\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)} =
\log \mathbf{S_2}\boldsymbol{x}(t, \gamma_1, \gamma_2) -
\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1),
\end{equation}
\ie a linear combination of the logarithms of first- and second-order coefficients.
As such, a non-linear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier.

@@

Owing to the scarcity of salient events in many natural scenes,
fine-grained classification is only made
possible by integrating signal information over a long temporal context.
Indeed, whereas a few seconds is often sufficient to recognize a speaker,
a musical instrument, or a genre, up to $30$ seconds may be required
to disambiguate two classes of auditory scenes with overlapping semantic content, \eg a train from a subway station or a quiet street from a park.
Depending on whether aggregation is performed in feature space or in decision space, the corresponding method is referred to as early or late integration.

A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by only storing their average values.
Going back to the definition of the scattering transform given in Section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity. With a slight abuse of notation, we denote the summarized features by
\begin{equation}
\mathbf{S}\boldsymbol{x}(\gamma) =
\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t\mbox{.}
\end{equation}

Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
\begin{equation}
\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big)\mbox{,}
\end{equation}
where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.

A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
On the contrary, once the scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$ has been unpadded of any boundary artifacts, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per $30$-second auditory scene, resulting in $10240$ instances overall.
However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.
Moreover, when applying late integration, the classifier optimizes accuracy at the frame level, not at the scene level, so its prediction might not yield optimal training error under the final evaluation metric.
In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.

@@

Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges \cite{Hsu2003}.

To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we substract the sample means $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ to $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and divide the result by the sample standard deviations $\sigma[\mathbf{S}\boldsymbol{X}] (\gamma)$.
The vectors $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.

>>>>

Before classification or clustering, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
In this section, we review three transformations which achieve these properties, namely logarithmic compression, temporal integration, and standardization.

@@

Many algorithms in pattern recognition, including nearest-neighbor classifiers and support vector machines (SVMs), tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation to all coefficients. In particular, we find that the logarithm performs particularly well in this respect.
Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.

@@

Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone--a source of spurious variability in scene classification.
Logarithmic compression linearizes this dependency, facilitating the construction of powerful invariants at the classifier stage.

For the task of musical genre recognition, second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ are sometimes normalized by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, since this decorrelates them from one another \cite{Anden2014}.
We note that taking the logarithm of such renormalized coefficients yields
\begin{equation}
\log \dfrac{\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)}{\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)} =
\log \mathbf{S_2}\boldsymbol{x}(t, \gamma_1, \gamma_2) -
\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1),
\end{equation}
\ie a linear combination of the logarithms of first- and second-order coefficients.
As such, a non-linear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier. \ja{Do we do this here or not? What is the point of this paragraph?}

@@

Owing to the scarcity of salient events in many natural scenes,
fine-grained classification is only made
possible by integrating signal information over a long temporal context.
Indeed, whereas a few seconds is often sufficient to recognize a speaker,
a musical instrument, or a genre, up to $30$ seconds may be required
to disambiguate two classes of auditory scenes with overlapping semantic content, \eg a train from a subway station or a quiet street from a park.
Depending on whether aggregation is performed in feature space or in decision space (the output of a classifier) the corresponding method is referred to as early or late integration.

A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by computing their average values over time.
In the definition of the scattering transform (see Section \ref{sec:scattering}), this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity. With a slight abuse of notation, we denote the summarized features by
\begin{equation}
\mathbf{S}\boldsymbol{x}(\gamma) =
\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t\mbox{.}
\end{equation}

Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$ obtained from some supervised model, such as a classifier. These are subsequently aggregated to produce a final decision
\begin{equation}
\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big)\mbox{,}
\end{equation}
where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}. \ja{Do we really sum probabilities? Seems like a strange thing to do.}

A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, which can increase the variance of the estimated model and run the risk of overfitting.
On the contrary, once the scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$ has been unpadded of any boundary artifacts, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per $30$-second auditory scene, resulting in $10240$ instances overall.
However, many of these instances may be silent or lack any salient properties of the class they are assigned to, which may increase the bias in the estimated model and run the risk of underfitting.
Moreover, when applying late integration, the classifier optimizes accuracy at the frame level, not at the scene level, so its prediction might not yield optimal training error under the final evaluation metric.
In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff. The appropriateness of either approach depends therefore on the nature of the signal that is classified.
We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority vote of the predictions given by an SVM.

@@

Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
